Le titre

### 2022-09-02 Manuscrit Thèse/chapter 4/Chapter 4.tex
bvfhkzdbvj

(mentionner l'article ici !).... The main motivation for this study was the potential to use NV-NV CR as a low field magnetometry protocol. Indeed, while NV-NV CR lines can be used to perform magnetometry with non zero fields [les russes], there are several advantages to use NV-NV CR as close to the zero magnetic field region, in particular the non dependence of the magnetic field orientation with respect to the crystal lattice. The behavior of

@@

@@

\subsection{Hyperfine coupling and inhomogeneous broadening}
\label{sec modif T2*}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/Comparaison_ESR_T2}
\caption{tata. Changer les T1 avec T1ph=5ms}
\label{ESR for T2*}
\end{figure}

@@

We should note that the current practice in magnetometry is to consider only of the three hyper-fine lines when they are resolved. This is because magnetometry protocol usually relies on a microwave field with a very well defined frequency, which can effectively select only one of the lines. In our case however, we have to consider the spectral overlap between NV centers and fluctuators, which have an additional broadening of $2\gamma_f \approx 6\ \rm MHz$ [REF] that completely obscures the hyper-fine structure. We therefore have to consider the full linewidth, even when the hyper-fine structure is resolved.

@@

Finally, \ref{ESR for T2*}-c) shows the linewidth of both samples for zero external magnetic field. For sample Sumi-2, the profile and linewidth is similar to the case of the purely transverse magnetic field, which is consistent with the fact that the electric field noise is stronger than the residual magnetic fields (earth magnetic field and hyper-fine interaction). For sample CVD-pink however, the electric field noise is smaller than the hyper-fine interaction, meaning that only the $\ket{m_I=0}$ states (the ones closest to the central dip) will be dominated by the electronic noise, and be in the electronic states $\ket{m_e=\pm}$. The $\ket{m_I=\pm 1}$ states are dominated by the hyper-fine field and are in the electronic $\ket{m_e=\pm 1}$ basis.

Overall, we are mostly interested in the linewidth of the ODMR lines for various magnetic field conditions. While these linewidth are quite significantly changed for sample CVD-pink, the change is less pronounced on sample Sumi-2. Most HPHT samples we used are from type 1B diamond, and behave similarly to sample Sumi-2. Since almost all measurements in this chapter will be conducted on HPHT samples, we will consider the modification of $T_2^*$ as a minor effect.

@@

\subsection{Experimental observations}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/scan_1x1x1x1}
\caption{tata. Changer les T1 avec T1ph=5ms}
\label{scan 1x1x1x1}
\end{figure}
We will start by showing that NV-NV CR behaves differently in the low magnetic field regime compared to the longitudinal field dominated regime which we studied in the last chapter.

@@

\section{NV-NV CR under purely transverse magnetic field}
We now want to experimentally discriminate these different contributions and to see how close the actual decay rates are to the predicted ones.

@@

Instead of an electric field, we will use a transverse magnetic field, which as described in sec. \ref{sec B transverse} leads to a spin Hamiltonian similar to the one where the transverse electric field dominates, or at least to similar eigenstates.
\subsection{Principle of the experiment}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/transis_transverse_field}
\caption{Simulation of the eigenstates of the NV Hamiltonian for $d_\perp E_perp=4\ \rm MHz$ and a purely transverse magnetic field $B_\perp$. (a) Eigenvalues for the spin Hamiltonian as a function of $B_\perp$. The three states are labeled $\ket{1}$, $\ket{2}$ and $\ket{3}$ (b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve)}
\label{eigenstates transverse field}
\end{figure}

@@

We have here quantitatively measured different depolarization contributions for a single class submitted to purely transverse magnetic field. We expect that the depolarization for low magnetic fields behave similarly, and that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
\begin{enumerate}
\item The lift of degeneracy between the four classes caused by the projection of the magnetic field on the different NV axis.
\item The double flips between the nearly resonant $\ket{+}$ and $\ket{-}$ states at low magnetic field.
\item The change of the Hamiltonian eigenbasis from $\{\ket{0}, \ket{\pm} \}$ in the electric field dominated region to $\{\ket{0}, \ket{\pm 1} \}$ in the magnetic field dominated region, and the incident change in $T_2^*$.
\end{enumerate}

\section{Introduction to ensemble NV magnetometry}
Before we explore the application of low field depolarization to magnetometry, we must discuss the general field of NV magnetometry.

@@

Ensemble NV centers center on the other hand offer sensitivities below $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of several hundreds of $\mu$m.

We will only focus here on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR. We will not try to give a complete overview of this field, but only to the elements close to the technique we present. For a general overview on microwave-based ensemble NV magnetometry, we recommend the excellent review from Barry et al \citep{barry2020sensitivity}.

@@

\subsection{Orientation-free magnetometry}
Every protocol described so far requires a precise knowledge  of the transition frequency between the $\ket{0}$ and $\ket{\pm 1}$ states versus the magnetic field. This knowledge implies that the direction of each NV axis in a frame of reference is known, because of the strong anisotropy in the NV Hamiltonian.

@@

Orientation-free magnetometry can also be performed for noise sensing, provided that that the noise is sufficiently broadband to directly excite the NV transitions regardless of their Larmor frequencies. The noise can be probed either by measuring the spin $T_1$ time \citep{kolkowitz2015probing, andersen2019electron} (which does not require a knowledge of the larmor frequency), or via the PL \citep{finco2021imaging}.

@@

We will first describe its operation and then characterize it and compare it to the previously mentioned NV magnetometry protocols.

@@

Fig. [A FAIRE] shows how the AC and DC sensing work for LFDM. For DC sensing, we want to use an oscillating field with an amplitude that matches the width of the PL dip $|\delta B| \approx \sigma$, in order to maximize the slope of the lock-in signal around $B_0=0$. For AC sensing, we want to use a static field $B_0=\sigma$ corresponding to the region where the lock-in signal is most sensitive to AC field.

@@

where $\tau=3\ r=\rm ms$ is the LIA low-pass filter time constant.

We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\delta V \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\delta V < 1\ \rm mV$). Laser amplitude fluctuations could also play a role but the noise seem to grow as the square root of the optical power and not linearly with it.

@@

\subsection{Temporal stability}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/Temporal_just_0B}
\caption{tata}
\label{temporal 0B}
\end{figure}

@@

\section{Comparison between LFDM and other magnetometry protocol}
We will now compare the LFDM performances we measured with the previously mentioned magnetometry protocols. We should start by mentioning that the best sensitivity achieved with the LFDM ($\sim 100 \rm{nT} / \sqrt{\rm Hz}$) is about $10^5$ times worse than the best NV ensemble sensitivity. This discrepancy however come in large part to the smaller sensor volume used in the characterization, and to a relatively low PL collection efficiency compared to the other experiments.

@@

\begin{table}[htbp]
\centering
\caption{Tutu}
 \label{table sensi volumiques}
\begin{tabular}{c|ccc}
\toprule
{} & Zhou et al. \citep{zhou2020quantum} (AC) & Barry et al.\citep{barry2016optical} (DC) & LFDM (DC) \\
\midrule
$\eta\ (\rm{nT} / \sqrt{\rm Hz})$ & 92&0.015& 116 \\
$V\ (\mu \rm{m}^3)$ &$8.1 \cdot 10^{-3}$ &$5.2 \cdot 10^{6}$& $3.3 \cdot 10^{3}$ \\
$\eta_v\ (\rm{nT}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2})$ &8.3&34&6700 \\
\bottomrule
\end{tabular}
\end{table}

@@

Table \ref{table sensi volumiques} reports the best AC and DC volume-normalized sensitivities found in the literature. The LFDM protocol performs $\sim 200$ times worse than the best DC magnetometer. It should be noted however that the PL collection in our measurement, where we use a confocal microscope using an objective with NA=0.65, is estimated to be $\sim 1 \%$ \citep{le2012efficient}, while the best collection schemes \citep{le2012efficient,wolf2015subpicotesla,wang2022picotesla, alsid2022solid} achieve a collection efficiency above 50 \%.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/compariason_0B_MW}
\caption{tata}
\label{comparaison 0B MW}
\end{figure}

@@

We should acknowledge that this microwave-based detection is not optimal: we are using a technique analogous to CW ODMR which is not as sensitive as Ramsey interference (although CW ODMR was also the technique used in \citep{barry2016optical}), we were limited in the microwave power (although $\Omega_{Rabi} \sim 1/T_2^*$ which is supposed to be optimal), and the sample used here had a relatively poor $T_2^* \approx 68\ \rm ns$. Nevertheless, this shows that the LFDM protocol is not that far from basic microwave-based magnetometry with the samples and the optical setup used here.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/substrat}
\caption{tata}
\label{substrat chelou}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/double_adamas}
\caption{tata}
\label{double dragon}
\end{figure}

@@

These two examples illustrate the complexity of finding the sample-dependent parameters that governs the PL contrasts in zero field. We have have seen a large variability in PL contrasts among sample grown in similar condition, which would indicate that those parameters are quite sensitive. If they were to be identified, samples with dedicated fabrication process could then substantially improve the LFDM sensitivity.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/Applications}
\caption{tata}
\label{applications}
\end{figure}

@@

Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always be within $\sim$ 1 $\mu$m of the sample. Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens. One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$ m red fluorescent diamond) are already available. I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ (\rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2})$ than for the 15 $\mu$m samples. This sensitivity is already better than some previous applications of wide field imaging at $20 6 \ (\rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}. Another potential advantage of LFDM is that a homogeneous magnetic field or microwave intensity along the sample is not required. Only a $\sim 10 \ \rm G$ AC magnetic field is required, and the sensitivity does not depend critically on the AC field intensity.

Another potential application for LFDM is shown in Fig. \ref{applications}-b. It consists in the detection of magnetic field gradient on top of a surface with a nano-diamond probe attached to a nano-resonator. The resonator motion transform the DC gradient in an AC field which can then be probed following the protocol described in Fig. [A faire]. This idea is based on the previous work with NV centers \citep{arcizet2011single} and magnetic resonance force microscopy (MRFM) \citep{rugar2004single}, and a very similar protocol as recently been developed with single NV centers \citep{huxter2022scanning}.

@@

Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations and does not require a homogeneous static magnetic field or microwave field.

>>>>

In the last chapter we saw how dense ensemble of NV centers have an intrinsic depolarization mechanism mediated by the NV-NV dipole interaction. In this chapter we will see how this feature can be exploited to perform magnetometry.

Using NV-NV CR for magnetometry in non-zero magnetic field has already been proposed \citep{akhmedzhanov2017microwave, akhmedzhanov2019magnetometry}. We will focus here on low to zero field magnetometry, which presents practical advantages as the relative orientation of the diamond with the magnetic field does not need to be controlled, but theoretical complications due to the complex physics of the NV centers in low magnetic field.

This chapter is constructed as follows: We will first describe the physics of NV centers under low or transverse magnetic field (the physics is similar in both cases), we will then show both theoretically and experimentally that the NV-NV CR is modified for low and transverse magnetic field. We will identify different mechanisms by which the relaxation is modified and we will try to quantify their relative importance.

We will then focus on the low field depolarization magnetometry (LFDM) protocol which exploits the previously studied NV-NV CR at low magnetic field. We will start by giving a short overview of NV ensemble magnetometry before characterizing the LFDM protocol and comparing it to other magnetometry protocols. Finally we will give some perspectives on the how to improve LFDM and on potential applications.

The results presented in this chapter are based in large part on the article \citep{pellet2022spin}.

@@

Parler du champs magnétique terrestre.

@@

\subsection{Hyperfine coupling and inhomogeneous broadening}
\label{sec modif T2*}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/Comparaison_ESR_T2}
\caption{ODMR lineshapes for longitudinal, transverse and low magnetic fields on samples CVD-pink and Sumi-2. a) ODMR spectrum of a single class of NV centers for a longitudinal magnetic field ($B_\parallel \sim 100\ \rm G$). The microwave frequency is given with respect to the ODMR line central frequency. b) ODMR spectrum of a single class for a purely transverse magnetic field ($B_\perp \sim 100\ \rm G$). c) ODMR of all 4 classes for 0 external magnetic field. d) Table reporting the full width at half maximum of each ODMR line.}
\label{ESR for T2*}
\end{figure}

@@

We should note that the usual practice in magnetometry is to consider only of the three hyper-fine lines when they are resolved. This is because magnetometry protocol usually relies on a microwave field with a very well defined frequency, which can effectively select only one of the lines. In our case however, we have to consider the spectral overlap between NV centers and fluctuators, which have an additional broadening of $2\gamma_f \approx 6\ \rm MHz$ [REF] that completely obscures the hyper-fine structure. We therefore have to consider the full linewidth, even when the hyper-fine structure is resolved.

@@

Finally, \ref{ESR for T2*}-c) shows the linewidth of both samples for zero external magnetic field. For sample Sumi-2, the profile and linewidth is similar to the case of the purely transverse magnetic field, which is consistent with the fact that the electric field noise is stronger than the residual magnetic fields (earth magnetic field and hyper-fine interaction). For sample CVD-pink however, the electric field noise is smaller than the hyper-fine interaction, meaning that only the $\ket{m_s=-1,m_I=+1}$ and $\ket{m_s=+1,m_I=-1}$ states (the ones closest to the central dip) will be dominated by the electronic noise, and be in the electronic states $\ket{m_e=\pm}$. The $\ket{m_I=\pm 1}$ states are dominated by the hyper-fine field and are in the electronic $\ket{m_e=\pm 1}$ basis.

The table in Fig. \ref{ESR for T2*}-d) reports the linewidths measured on both sample for the various magnetic field configurations. We can see that the linewidth of sample CVD-pink varies by more than a factor of 2, while the ones from sample Sumi-2 varied by less than 20 \%. We should note that  almost all samples used in this chapter are HPHT type 1b diamond, similar to sample Sumi-2 and unlike sample CVD-pink. We should therefore expect the modification of $T_2^*$ to have a moderate impact on the depolarization dynamics.

\medskip

To summarize this section, we have seen that (for the samples used in this chapter) the NV centers spin Hamiltonian is dominated by transverse electric field for low external magnetic field, and that it causes a change of the Hamiltonian eigenstates from $\{ \ket{0}, \ket{\pm 1} \}$ to $\{ \ket{0}, \ket{\pm} \}$. We have seen that the Hamiltonian has the same eigenstates for a purely transverse magnetic field, and finally we have seen that this change of eigenstates occasion a modification of the spins $T_2^*$, although this modification is slight for samples charged in impurities.

@@

\subsection{Experimental observations}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/scan_1x1x1x1}
\caption{Low field depolarization on sample ADM-150-1 for $\mathbf{B}$ in an arbitrary direction. a) ODMR spectrum for B=0. b) ODMR spectrum for B=60 G. c) PL contrast as a function of the external magnetic field. d) $1/T_1^{\rm dd}$ as a function of the external magnetic field. A $T_1$ measurement was recorded for each magnetic field value and fitted accorded to the protocol described in [REF] with $T_1^{\rm ph}=5\ \rm ms$}
\label{scan 1x1x1x1}
\end{figure}
We will start by showing that NV-NV CR behaves differently in the low magnetic field regime compared to the longitudinal field dominated regime which we studied in the last chapter.

@@

\section{NV-NV CR under purely transverse magnetic field}
We have seen in the last part that there was a spontaneous spin depolarization specifically in zero external magnetic field, and we have identified three potential mechanisms which are the double-flips, the change in the Hamiltonian eigenstates and the modification of $T_2^*$. We now want to experimentally discriminate the contribution of each of these mechanisms and to see how close the actual decay rates are to the predicted ones.

@@

Instead of an electric field, we will use a transverse magnetic field, which as described in sec. \ref{sec B transverse} leads to a spin Hamiltonian similar to the one where the transverse electric field dominates, or at least to similar eigenstates.

\subsection{Principle of the experiment}
\begin{figure}[h]
\centering
\includegraphics[width=\textwidth]{Figures/transis_transverse_field}
\caption{Simulation of the eigenstates of the NV Hamiltonian for $d_\perp E_perp=4\ \rm MHz$ and a purely transverse magnetic field $B_\perp$. (a) Eigenvalues for the spin Hamiltonian as a function of $B_\perp$. The three states are labeled $\ket{1}$, $\ket{2}$ and $\ket{3}$ (b) Splitting between the $\ket{2}$ and $\ket{3}$ states (blue curve) and closeness factor $\abs{\bra{3}\ket{+}}^2$ between the $\ket{3}$ and $\ket{+}$ (red curve)}
\label{eigenstates transverse field}
\end{figure}

@@

We have here quantitatively measured different depolarization contributions for a single class submitted to purely transverse magnetic field, and we expect that the depolarization for low magnetic fields behave similarly. We therefore claim that the difference in NV polarization between the $\mathbf{B}=0$ region and $\mathbf{B} \neq 0$ region is caused by (in order of importance):
\begin{enumerate}
\item The lift of degeneracy between the four classes caused by the projection of the magnetic field on the different NV axis.
\item The double flips between the nearly resonant $\ket{+}$ and $\ket{-}$ states at low magnetic field.
\item The change of the Hamiltonian eigenbasis from $\{\ket{0}, \ket{\pm} \}$ in the electric field dominated region to $\{\ket{0}, \ket{\pm 1} \}$ in the magnetic field dominated region, and the incident change in $T_2^*$.
\end{enumerate}

\section{Introduction to ensemble NV magnetometry}

Before we explore the application of low field depolarization to magnetometry, we must discuss the general field of NV magnetometry.

@@

Ensemble NV centers center on the other hand offer sensitivities below $1\ \rm pT / \sqrt{\rm Hz}$ \citep{wolf2015subpicotesla}, at the cost of working with samples of several hundreds of up to several mm.

We will only focus here on NV ensemble magnetometry since we want to give elements of comparisons with our technique based on NV-NV CR. We will not try to give a complete overview of this field, but only of the elements close to the technique we present. For a general overview on NV magnetometry, we recommend the excellent following reviews  \citep{rondin2014magnetometry, degen2017quantum, barry2020sensitivity}.

@@

\subsection{Orientation-free magnetometry}
Every protocol described so far requires a precise knowledge  of the Larmor frequency between the $\ket{0}$ and $\ket{\pm 1}$ states as a function of the magnetic field. This knowledge implies that the direction of each NV axis in the frame of the magnetic field is known, because of the strong anisotropy in the NV Hamiltonian.

@@

Orientation-free magnetometry can also be performed for noise sensing, provided that that the noise is sufficiently broadband to directly excite the NV transitions regardless of their Larmor frequencies. The noise can be probed either by measuring the spin $T_1$ time \citep{kolkowitz2015probing, andersen2019electron} (which does not require a knowledge of the Larmor frequency), or via the PL \citep{finco2021imaging}.

@@

We will first describe its operation and characterize it before comparing it to the previously mentioned NV magnetometry protocols.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.9\textwidth]{Figures/AC_and_DC_sensing}
\caption{DC and AC LFDM sensing protocols on sample ADM-15-5. The magnetic field is scanned in an arbitrary direction. a) DC sensing: $B_0$ is scanned between $-$80 and +80 G while an AC field of amplitude 10 G is used for the lock-in detection. b) AC sensing: $B_0$ scans are performed for 5 different $|\tilde B_{\rm AC}|$ values between 2 and 10 G. The best AC sensitivity is achieved for $B_0\approx 5\ \rm G$.}
\label{AC and DC sensing}
\end{figure}

Fig. \ref{AC and DC sensing} shows how the AC and DC sensing work for LFDM. For DC sensing, we want to use an oscillating field with an amplitude that matches the width of the PL dip $|\delta B| \approx \sigma$, in order to maximize the slope of the lock-in signal around $B_0=0$. For AC sensing, we want to use a static field $B_0\approx \sigma/2$ corresponding to the region where the lock-in signal is most sensitive to AC field.

@@

where $\tau=3\ \rm ms$ is the LIA low-pass filter time constant.

We estimate that the sensitivity is mostly shot-noise limited because the same LIA voltage noise ($\delta V \approx 35\ \rm mV$) was observed on a magnetic insensitive region (for $B_0 \approx 100\ \rm G$), and is far above the electrical noise measured when no optical signal is sent to the photodiode ($\delta V < 1\ \rm mV$). Laser fluctuations could also play a role but the noise seem to grow as the square root of the optical power and not linearly with it.

@@

\subsection{Temporal stability}
\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/Temporal_just_0B}
\caption{Temporal stability of DC-LFDM as measured on sample ADM-15-5. a) Temporal trace of the LIA signal for $B_0 \approx 0$. The Lock-in signal was converted to a measured magnetic field with the previously described protocol. b) Estimated signal to noise ratio for a signal of 100 nT as a function of the total acquisition time $T$. The noise value is computed by averaging the variance of the temporal trace over each time duration $T$. A curve $y=0.5\, T^{1/2}$ is added for comparison}
\label{temporal 0B}
\end{figure}

@@

\section{Comparison between LFDM and other magnetometry protocol}
Now that we have characterized the LFDM protocol, we can compare its performances with the previously mentioned magnetometry protocols. We should start by mentioning that the best sensitivity achieved with the LFDM ($\sim 100 \rm{nT} / \sqrt{\rm Hz}$) is about $10^5$ times worse than the best NV ensemble sensitivity. This discrepancy however come in large part to the smaller sensor volume used in the characterization, and to a relatively low PL collection efficiency compared to the other experiments.

@@

\begin{table}[htbp]
\centering
\caption{Comparison of the volume-normalized sensitivity of the best AC and DC protocols with LFDM}
 \label{table sensi volumiques}
\begin{tabular}{c|ccc}
\toprule
{} & Zhou et al. \citep{zhou2020quantum} (AC) & Barry et al.\citep{barry2016optical} (DC) & LFDM (DC) \\
\midrule
$\eta\ (\rm{nT} / \sqrt{\rm Hz})$ & 92&0.015& 116 \\
$V\ (\mu \rm{m}^3)$ &$8.1 \cdot 10^{-3}$ &$5.2 \cdot 10^{6}$& $3.3 \cdot 10^{3}$ \\
$\eta_v\ (\rm{nT}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2})$ &8.3&34&6700 \\
\bottomrule
\end{tabular}
\end{table}

@@

Table \ref{table sensi volumiques} reports the best AC and DC volume-normalized sensitivities found in the literature. The LFDM protocol performs $\sim 200$ times worse than the best DC magnetometer. It should be noted however that the PL collection in our measurement, where we use a confocal microscope with a numerical aperture NA=0.65, is estimated to be $\sim 1 \%$ \citep{le2012efficient}, while the best collection schemes \citep{le2012efficient,wolf2015subpicotesla,wang2022picotesla, alsid2022solid} achieve a collection efficiency above 50 \%.

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/compariason_0B_MW}
\caption{Comparison between LFDM and microwave-based magnetometry on the same experimental setup. The sample used is ADM-15-5 and the magnetic field is aligned close to a [111] axis. a) Schematics of the experiment: the setup is the same as the one described in Fig. \ref{setup magneto} with the addition of a continuous microwave tone at the frequency $\nu=2600\ \rm MHz$. b) LIA signal as $B_0$ is scanned from $-$150 to +150 G. The two lines at $\pm 100\ \rm G$ are caused by the microwave. c) Temporal trace of the measured B field for $B_0 \approx 0$ (on the middle of the zero field slope) and $B_0 \approx 100\ \rm G$ (on the middle of the microwave slope). The orange curve is shifted by $\approx 100\ \rm G$ to sit on the same scale as the blue curve. d) Estimated SNR fo a signal of 100 nT for $B_0 \approx 0$ and $B_0 \approx 100\ \rm G$ as a function of the total acquisition time}
\label{comparaison 0B MW}
\end{figure}

@@

We should acknowledge that this microwave-based detection is not optimal: we are using a technique analogous to CW ODMR which is not as sensitive as Ramsey interference (although CW ODMR was also the technique used in \citep{barry2016optical}), we were limited in the available microwave power (although $\Omega_{Rabi} \sim 1/T_2^*$ which is supposed to be optimal), and the sample used here had a relatively poor $T_2^* \approx 68\ \rm ns$. Nevertheless, this shows that the LFDM protocol is not that far from basic microwave-based magnetometry with the samples and the optical setup used here.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/substrat}
\caption{Measurements on sample SBST-C for two different regions A and B. a) PL contrast as a function of a magnetic field scanned in an aribitrary direction (same direction for both regions). b) PL contrast as a function of a magnetic field scanned in the [100] direction. c) $T_1$ measurement on a single class when $B\neq 0$. d) $T_1$ measurement on all  four classes when $B=0$ }
\label{substrat chelou}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/double_adamas}
\caption{Same measurements as Fig. \ref{substrat chelou} on two distinct samples ADM-15-4 and ADM-15-1. Every parameters (laser power,optical setup , AC magnetic field) was kept identical for the two samples.}
\label{double dragon}
\end{figure}

@@

These two examples illustrate the complexity of finding the sample-dependent parameters that governs the PL contrasts in zero field. We have noticed a large variability in PL contrasts among sample grown in similar condition, which would indicate that those parameters are quite sensitive. If these parameters were to be identified, samples with dedicated fabrication process could then substantially improve the LFDM sensitivity.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=\textwidth]{Figures/Applications}
\caption{Schematics of two potential application for LFDM. The working of each application is detailed in the main text}
\label{applications}
\end{figure}

@@

Instead we propose to deposit a thin layer of 1 $\mu$m diamonds on top of the sample in order to always be within $\sim$ 1 $\mu$m of the sample. Going below 1 $\mu$m may not prove useful as the spatial resolution is ultimately limited by the diffraction limit of the lens. One advantage of this method is that commercially available samples (Adamas Nano 1 $\mu$ m red fluorescent diamond) are well suited for this task. I have measured the sensitivity on some of these samples and found the same volumetric sensitivity $\eta_v \sim 6 \ \rm{\mu T}\mu \rm{m}^{3/2} \rm{Hz}^{-1/2}$ than for the 15 $\mu$m samples. This sensitivity is already better than some previous applications of wide field imaging at $20  \ \rm{\mu T}\mu \rm{m} \rm{Hz}^{-1/2}$ \citep{glenn2017micrometer}. Another potential advantage of LFDM is that a homogeneous magnetic field or microwave intensity along the sample is not required. Only a $\sim 10 \ \rm G$ AC magnetic field is required, and the sensitivity does not depend critically on the AC field homogeneity.

Another potential application for LFDM is shown in Fig. \ref{applications}-b. It consists in the detection of magnetic field gradient on top of a surface thanks to a nano-diamond probe attached to a nano-resonator. The resonator motion transform the DC gradient in an AC field which can then be probed following the protocol described in Fig. \ref{AC and DC sensing}-b). This idea is based on the previous work with NV centers \citep{arcizet2011single} and magnetic resonance force microscopy (MRFM) \citep{rugar2004single}. A very similar protocol was recently developed with microwave based AC sensing \citep{huxter2022scanning}.

@@

Finally, LFDM could be a good candidate to be scaled up since the protocol is not sensitive to strain or thermal variations and does not require a homogeneous magnetic field or microwave field along the sample.
